---
title: "Acessing fish ecological parameters, traits or indicators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Acessing fish ecological parameters, traits or indicators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(httr2)
library(fwtraits)
```

**Step 1: Get the before you start page for instructions on how to access the database**.

```{r beforeyoure start procedure}
#fw_be4ustart()
```

**Step 2: Get authentication token through the API key**.

To publicly use the API key the following steps based on https://httr2.r-lib.org/articles/wrapping-apis.html are used

a) Make a secret key using **`secret_make_key()`** and save it R-Environment using 
**`usethis::edit_r_environ()`**. Please install usethis package in case you don't have it installed. In the R environment save it with an identifier such as **FWTRAITS = 'keymadein_secret_make_key function'**.

b) Encrypt the API key obtained from the database administrators or during registration using 
**`secret_encrypt()`**.

c) From b) a scrambled key is generated which will be used in **`secret_decrypt()`**


```{r generate token}

#
#encrypted token for my api key

#sessionkey <- secret_make_key()

#edit this page with usethis::edit_r_environ()

#enc_api <- secret_encrypt(x = 'apikey', key = sessionkey)

enc_api <- "p6-9gAvYXveyC_B_0hTzyYl5FwLRwZEPD-ZE9Y4KrIBstgNc8K2Mr4u6t39LGZ2E1m9SOw"

#the FWTRAITS_KEY is the unlock key saved in my local environment
#check https://httr2.r-lib.org/articles/wrapping-apis.html for more information
apikey <- httr2::secret_decrypt(encrypted = enc_api, key = 'FWTRAITS_KEY')


#download fish catchment region data

apikeydecrypted <- fw_loadapikey(test = TRUE, encrytedkey = enc_api,
                              fwtraitskey =  'FWTRAITS_KEY')

tokendata <- fw_token(key= apikeydecrypted, seed = 1234)

```

**Access the fish traits data**

```{r access species ecological traits}

#One species 

migration <- fw_fetchdata(data = "Abramis brama", organismgroup = 'fi',
                          ecoparams = 'migration', 
                          token = tokendata)

data("speciesdata") #data archived in fwtraits

fishtraits <- fw_fetchdata(data = speciesdata, organismgroup = 'fi', token = tokendata,
                           spcol = 'scientificName', 
                           ecoparams = c('migration', 'catchment region'),
                           parallel = TRUE, cores = 3, sanitize = TRUE)

```

**Visualize**

```{r viszualize outputs}

fw_visualize(foutput = fishtraits, params = 'migration')

fw_visualize(foutput = fishtraits)

```


**References**
* Schmidt-Kloiber, A., & Hering, D. (2015). www.freshwaterecology.info - An online tool that unifies, standardises and codifies more than 20,000 European freshwater organisms and their ecological preferences. Ecological Indicators, 53, 271-282. https://doi.org/10.1016/j.ecolind.2015.02.007





