---
title: "Acessing macroinvertebrates data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Acessing macroinvertebrates data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(httr2)
library(fwtraits)
```

**Step 1: Get the before you start page for instructions on how to access the database**.

```{r beforeyoure start procedure}
#fw_be4ustart()
```

**Step 2: Get authentication token through the API key**.

To publically use the API key the following steps based on https://httr2.r-lib.org/articles/wrapping-apis.html are used

a) Make a secret key using **`secret_make_key()`** and save it R-Environment using 
**`usethis::edit_r_environ()`**. Please install usethis package in case you don't have it installed. In the R environment save it with an identifier such as **FWTRAITS = 'keymadein_secret_make_key function'**.

b) Encrypt the API key obtained from the database administrators or during registration using 
**`secret_encrypt()`**.

c) From b) a scrambled key is generated which will be used in **`secret_decrypt()`**


```{r generate token}

#
#encrypted token for my api key

#sessionkey <- secret_make_key()

#edit this page with usethis::edit_r_environ()

#enc_api <- secret_encrypt(x = 'apikey', key = sessionkey)

enc_api <- "p6-9gAvYXveyC_B_0hTzyYl5FwLRwZEPD-ZE9Y4KrIBstgNc8K2Mr4u6t39LGZ2E1m9SOw"

#the FWTRAITS_KEY is the unlock key saved in my local environment
#check https://httr2.r-lib.org/articles/wrapping-apis.html for more information
apikey <- httr2::secret_decrypt(encrypted = enc_api, key = 'FWTRAITS_KEY')


#download fish catchment region data

apikeydecrypted <- fw_loadapikey(test = TRUE, encrytedkey = enc_api,
                              fwtraitskey =  'FWTRAITS_KEY')

tokendata <- fw_token(key= apikeydecrypted, seed = 1234)

```

**Access the fish traits data**

**NOTE**: Before accessing the species ecological preferences, the all data will first be downloaded even if one species is indicated. But after, the data will be loaded from memory for the proceeding queries. 

```{r access species ecological traits}

#One species 

mmdata_traits <- fw_fetchdata(data =  "Margaritifera margaritifera",
                            organismgroup = 'mi',
                            ecoparams = c('ecoregional distribution sensu',
                                          'tolerance score',
                                            'sensitive speces',
                                          'stream zonation preference',
                                          'locomotion type'),
                            token = tokendata,
                            parallel = TRUE)

#for more than one species 

specieslist <- c("Margaritifera margaritifera",
                 "Pseudunio auricularius",
                 "Musculium lacustre",
                 "Musculium transversum",
                 "Parastenocaris germanica")

multspp <- fw_fetchdata(data =  specieslist,
                            organismgroup = 'mi',
                            ecoparams = c('ecoregional distribution sensu',
                                          'tolerance score',
                                            'sensitive speces',
                                          'stream zonation preference',
                                          'locomotion type'
                                            ),
                            token = tokendata,
                        parallel = TRUE, cores = 4,
                        sanitize = TRUE)

```

**Visualize**

```{r viszualize outputs}

#fw_visualize(foutput = mmdata_traits)# not necessary since its one species

fw_visualize(foutput = multspp)

```

**References**
* Schmidt-Kloiber, A., & Hering, D. (2015). www.freshwaterecology.info - An online tool that unifies, standardises and codifies more than 20,000 European freshwater organisms and their ecological preferences. Ecological Indicators, 53, 271-282. https://doi.org/10.1016/j.ecolind.2015.02.007

