---
title: "Acessing macroinvertebrates data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Acessing macroinvertebrates data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(httr2)
library(fwtraits)
```

###Step 1: Get the before you start page for instructions on how to access the database.

```{r beforeyoure start procedure}
#fw_be4ustart()
```

###Step 2: Get authentication token through the API key.

To publically use the API key the following steps based on https://httr2.r-lib.org/articles/wrapping-apis.html are used

a) Make a secret key using **`secret_make_key()`** and save it R-Environment using 
**`usethis::edit_r_environ()`**. Please install usethis package in case you don't have it installed. In the R environment save it with an identifier such as **FWTRAITS = 'keymadein_secret_make_key function'**.

b) Encrypt the API key obtained from the database administrators or during registration using 
**`secret_encrypt()`**.

c) From b) a scrambled key is generated which will be used in **`secret_decrypt()`**


```{r generate token}

#
#encrypted token for my api key

#sessionkey <- secret_make_key()

#edit this page with usethis::edit_r_environ()

#enc_api <- secret_encrypt(x = 'apikey', key = sessionkey)

enc_api <- "p6-9gAvYXveyC_B_0hTzyYl5FwLRwZEPD-ZE9Y4KrIBstgNc8K2Mr4u6t39LGZ2E1m9SOw"

#the FWTRAITS_KEY is the unlock key saved in my local environment
#check https://httr2.r-lib.org/articles/wrapping-apis.html for more information

#download fish catchment region data

apikeydecrypted <- fw_loadapikey(test = TRUE, sacrambled_apikey = enc_api,
                              fwtraitskey =  'FWTRAITS_KEY')

tokendata <- fw_token(key= apikeydecrypted, seed = 1234)

```

###Step3: Access the macroinvertebrates traits data

* **NOTE**: Before accessing the species ecological preferences, all the indicates traits data will first be downloaded even if one species is indicated. But after, the data will be loaded from computer memory for the proceeding queries. 

```{r access species ecological traits}

#One species 

mmdata_traits <- fw_fetchdata(data =  "Margaritifera margaritifera",
                            organismgroup = 'mi',
                            ecoparams = c('stream zonation preference','locomotion type'),
                            token = tokendata,
                            parallel = TRUE)

#for more than one species in a list 

specieslist <- c("Margaritifera margaritifera",
                 "Pseudunio auricularius",
                 "Musculium lacustre",
                 "Musculium transversum",
                 "Parastenocaris germanica")

multspp <- fw_fetchdata(data =  specieslist,
                            organismgroup = 'mi',
                            ecoparams = c('stream zonation preference',
                                          'locomotion type'),
                            token = tokendata,
                        parallel = TRUE, cores = 2,
                        sanitize = TRUE)

#species in a data frame 

macrodf <- data.frame(organismgroup = rep('mi', 5),
                      species = c("Margaritifera margaritifera",
                 "Pseudunio auricularius",
                 "Musculium lacustre",
                 "Musculium transversum",
                 "Parastenocaris germanica"))

multspp_df <- fw_fetchdata(data =  macrodf,
                            organismgroup = 'mi',
                           spcol = 'species',
                            ecoparams = c('stream zonation preference',
                                          'locomotion type'),
                            token = tokendata,
                        parallel = TRUE, cores = 4,
                        sanitize = TRUE)

```

###Step4. Visualize extracted outputs

```{r viszualize outputs, fig.width = 6, fig.height= 5, fig.align='center'}

#fw_visualize(foutput = mmdata_traits)# not necessary since its one species

fw_visualize(foutput = multspp)

fw_visualize(foutput = multspp_df)

```

**References**
* Schmidt-Kloiber, A., & Hering, D. (2015). www.freshwaterecology.info - An online tool that unifies, standardises and codifies more than 20,000 European freshwater organisms and their ecological preferences. Ecological Indicators, 53, 271-282. https://doi.org/10.1016/j.ecolind.2015.02.007

