---
title: "Accessing species ecological preferences for multiple organism groups."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessing species ecological preferences for multiple organism groups.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r setup}
library(httr2)
library(fwtraits)
```

**Step 1: Get the before you start page for instructions on how to access the database**.

```{r beforeyoure start procedure}
#fw_be4ustart()
```

**Step 2: Get authentication token through the API key**.

To publically use the API key the following steps based on https://httr2.r-lib.org/articles/wrapping-apis.html are used

a) Make a secret key using **`secret_make_key()`** and save it R-Environment using 
**`usethis::edit_r_environ()`**. Please install usethis package in case you don't have it installed. In the R environment save it with an identifier such as **FWTRAITS = 'keymadein_secret_make_key function'**.

b) Encrypt the API key obtained from the database administrators or during registration using 
**`secret_encrypt()`**.

c) From b) a scrambled key is generated which will be used in **`secret_decrypt()`**


```{r generate token}

#
#encrypted token for my api key

#sessionkey <- secret_make_key()

#edit this page with usethis::edit_r_environ()

#enc_api <- secret_encrypt(x = 'apikey', key = sessionkey)

enc_api <- "p6-9gAvYXveyC_B_0hTzyYl5FwLRwZEPD-ZE9Y4KrIBstgNc8K2Mr4u6t39LGZ2E1m9SOw"


#download fish catchment region data

apikeydecrypted <- fw_loadapikey(test = TRUE, sacrambled_apikey = enc_api,
                              fwtraitskey =  'FWTRAITS_KEY')

tokendata <- fw_token(key= apikeydecrypted, seed = 1234)

```

**Access the multiple organism groups ecological preferences data**

**NOTE**: Before accessing the species ecological preferences, the all data will first be downloaded even if one species is indicated. But after, the data will be loaded from memory for the proceeding queries. 

```{r access species ecological traits}

#list of species data

specieslist <- list(fi= c('Abramis brama', 'Salmo truta fario', "hUcho hucho"),
                    mi= c("Margaritifera margaritifera",
                          "Pseudunio auricularius",
                          "Musculium lacustre",
                          "Musculium transversum",
                          "Parastenocaris germanica"),
                    pb = c('Bangia atropurpurea', 'Aphanocapsa fonticola',
                                  'Aphanocapsa rivularis', 
                                  'Chamaesiphon incrustans',
                                  'Cyclostephanos dubius'),
                    pp = c("Achnanthidium exiguum","Achnanthidium exilis",
                           "Achnanthidium minutissima", "Amphora copulata",
                           "Amphora holsatica"),
                    di = c("Achnanthes acares","Achnanthes acus",
                           "Achnanthes alpestris","Achnanthes angustata",
                           "Achnanthes arctica","Achnanthes arenaria",
                           "Achnanthes atacamae","Achnanthes alpestris"),
                    mp = c("Amblystegium humile","Calliergon cordifolium",
                           "Scorpidium revolvens", "Philonotis calcarea"))

ecologicaltraits <- list(fi= c('threat austria', 'migration', "migartion fibs",
                               'oxygen tolerance'),
                    mi = c('ecoregional distribution sensu',
                                          'tolerance score',
                                            'sensitive speces',
                                          'stream zonation preference',
                                          'locomotion type'),
                    pb = c('life form', 'water type',"substrate preference",
                           'geochemical imprint'),
                    pp = c("life form","phytoplankton guild phytoloss",
                            "mode nutrition",
                           "seasonal occurrence rivers", "trophic focus",
                           "stenoecy factor"),
                    di = c("trophy bayern","trophy tdi",
                           "pollution ips","salinity denys"),
                    mp = c("plant life span","life form systema",
                           "life form raunkiaer", "growth form wiegleb",
                           "leaf persistance"))

multigroupsdata <- fw_fetchdata(data =  specieslist,
                            organismgroup = c('mi',"pp","di","mp","fi", 'pb'),
                            ecoparams = ecologicaltraits,
                            token = tokendata,
                            parallel = TRUE, cores = 2,
                            na.rm = FALSE,
                            sanitize = TRUE)

```


**Visualize**

```{r viszualize outputs, fig.width = 6, fig.height= 5, fig.align='center'}

fw_visualize(multigroupsdata, group = 'phytobentho')

fw_visualize(foutput = multigroupsdata, group = 'fishes')

```

**References**
* Schmidt-Kloiber, A., & Hering, D. (2015). www.freshwaterecology.info - An online tool that unifies, standardises and codifies more than 20,000 European freshwater organisms and their ecological preferences. Ecological Indicators, 53, 271-282. https://doi.org/10.1016/j.ecolind.2015.02.007
