---
title: "Extract traits for phytoplankton species ecological preferences."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extract traits for phytoplankton species ecological preferences.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(httr2)
library(fwtraits)
```

**Step 1: Get the before you start page for instructions on how to access the database**.

```{r beforeyoure start procedure}
#fw_be4ustart()
```

**Step 2: Get authentication token through the API key**.

To publicly use the API key the following steps based on https://httr2.r-lib.org/articles/wrapping-apis.html are used

a) Make a secret key using **`secret_make_key()`** and save it R-Environment using 
**`usethis::edit_r_environ()`**. Please install usethis package in case you don't have it installed. In the R environment save it with an identifier such as **FWTRAITS = 'keymadein_secret_make_key function'**.

b) Encrypt the API key obtained from the database administrators or during registration using 
**`secret_encrypt()`**.

c) From b) a scrambled key is generated which will be used in **`secret_decrypt()`**


```{r generate token}

#
#encrypted token for my api key

#sessionkey <- secret_make_key()

#edit this page with usethis::edit_r_environ()

#enc_api <- secret_encrypt(x = 'apikey', key = sessionkey)

enc_api <- "p6-9gAvYXveyC_B_0hTzyYl5FwLRwZEPD-ZE9Y4KrIBstgNc8K2Mr4u6t39LGZ2E1m9SOw"

#the FWTRAITS_KEY is the unlock key saved in my local environment
#check https://httr2.r-lib.org/articles/wrapping-apis.html for more information
apikey <- httr2::secret_decrypt(encrypted = enc_api, key = 'FWTRAITS_KEY')


#download fish catchment region data

apikeydecrypted <- fw_loadapikey(test = TRUE, encrytedkey = enc_api,
                              fwtraitskey =  'FWTRAITS_KEY')

tokendata <- fw_token(key= apikeydecrypted, seed = 1234)

```

**Access the fish traits data**

```{r access species ecological traits}

#One species 

getdf <- fw_searchdata(organismgroup = 'pp', token = tokendata)

d1 <- getdf[[1]][[10]] %>% mutate(speciesname = paste0(Genus, ' ', Species))#[2321,9]

phty <- fw_fetchdata(data = d1, 
                     organismgroup = 'pp',
                     spcol = 'speciesname',
                     warn = TRUE,
                     token = tokendata, 
                     na.rm = FALSE)




sund <- d1 %>% group_by(TaxaGroup) %>% summarise(lt= length(TaxaGroup))

stenoecyfactor <- fw_searchdata(organismgroup = 'pp', ecoparams = 'stenoecyfactor',
                       token = tokendata)
d2 <- stenoecyfactor[[1]][[2]][[9]]

lakehabitat <- fw_fetchdata(data = "Anabaena bergii", 
                               organismgroup = 'pp',
                               ecoparams = 'trophic focus', 
                               warn = TRUE,
                               token = tokendata, 
                            na.rm = FALSE)




sppe <- fw_fetchdata(data = "Achnanthidium exiguum", 
                                organismgroup = 'pp',
                                ecoparams = ' habitat - river/lake', 
                               warn = TRUE,
                               token = tokendata)

pp = c("life form","phytoplankton guild phytoloss",
                            "mode nutrition",
                           "seasonal occurrence rivers", "trophic focus",
                           "stenoecy factor")

pp = c("Achnanthidium exiguum","Achnanthidium exilis",
                           "Achnanthidium minutissima", "Amphora copulata",
                           "Amphora holsatica")

```


