---
title: "How to handle the API key in public and private codes."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to handle the API key in public and private codes.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fwtraits)

```

## Steps in securing the API key.

* STEP 1: Acquire the API key, the user can follow the steps elucidated in **`fw_be4ustart()`** function.

* STEP 2: Edit the .Renviron file by running **`fw_setapikey()`**. 

* STEP 3: If the .Renviron window opens, please type in
**API_KEY = "9f7be425-e099-11ee--.............."** 

* STEP 4: Save and restart the R session.

* STEP 5: The key will be ready for use and will not required to be provided anywhere within the codes thereafter.

* STEPT 6: When you test with the codes, a welcome message will be printed before data is retrieved.


### Test if the API key is well saved

```{r}


migration <- fw_fetchdata(data = 'Abramis brama', 
                         organismgroup = 'fi',
                         ecoparams = 'migration',
                         cachefolder = 'cache', 
                         warn = TRUE,
                         inform = TRUE,
                         details = TRUE)#the species spelling is checked

#or use the fw_keyload() function: which returns a congratulation message

```


## 3. References

1. Wickham H (2024). _httr2: Perform HTTP Requests and Process the Responses_. R package version 1.0.3,<https://CRAN.R-project.org/package=httr2>

2. https://httr2.r-lib.org/articles/wrapping-apis.html












